---
title: "Registered cats in Switzerland"
subtitle: "2016-01-01 to 2023-12-31"
author: "Jadey Ryan"
date-modified: last-modified
format: 
  html: default
embed-resources: true
execute: 
  echo: false
---

```{r}
#| label: setup
#| include: false

# Attach packages
library(readr)
library(here)
library(dplyr)
library(ggplot2)
library(stringr)

# Load data
pets <- read_rds(here("demos/data", "pets.RDS"))
```

## Breeds of cats in Switzerland

```{r}
#| label: breeds-tbl

breeds <- pets |> 
  filter(pet_type == "cats")

# Top 5 per date
breeds_top_5 <- breeds |>
  slice_max(order_by = count, n = 5, by = date)

# Top 5 overall (average across all dates)
breeds_avg <- breeds_top_5 |>
  summarize(
    "Average Number" = format(
      mean(count),
      big.mark = ",",
      digits = 1,
      justify = "right",
      scientific = FALSE
    ),
    .by = "breed"
  ) |>
  rename("Breed" = breed)
```

The breed with the largest average population in Switzerland from 2016 - 2023 was the **`r breeds_avg$Breed[[1]]`**.

```{r}
#| label: breeds-plot

breeds_top_5 |>
  ggplot(
    aes(x = date, y = count, color = breed)
  ) +
  geom_line(linewidth = 1) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(
    labels = \(x) format(x, big.mark = ",", scientific = FALSE)
  ) +
  labs(
    title = "Evolution by breed",
    subtitle = str_wrap(
      paste(
        "Number of registered and living",
        "cats",
        "by breed. The evolution of the largest populations is presented."
      ),
      80
    ),
    color = NULL,
    caption = "Source: Identitas AG Animal Statistics"
  ) +
  xlab(NULL) +
  ylab(NULL) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  theme(
    plot.caption.position = "plot",
    legend.position = "top",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

## My favorite breed: Snowshoe

```{r}
#| label: fave-breed
fave_breed <- breeds |>
  filter(breed == "Snowshoe")
```

```{r}
#| label: fave-breed-plot

fave_breed |>
  ggplot(
    aes(x = date, y = count)
  ) +
  geom_line(linewidth = 1) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  scale_y_continuous(
    labels = \(x) format(x, big.mark = ",", scientific = FALSE)
  ) +
  labs(
    title = paste("Snowshoe", "population"),
    caption = "Source: Identitas AG Animal Statistics"
  ) +
  xlab(NULL) +
  ylab(NULL) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  theme(
    plot.caption.position = "plot",
    legend.position = "top",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

## Data source

[Identitas AG](https://tierstatistik.identitas.ch/en/index.html) is the leading service provider for animal health, welfare and food safety in Switzerland. 

Dogs are obliged to be registered for licensing purposes at a local government level and this data is kept in the Amicus database. Other pets, such as cats, rodents, birds or fish, may be registered and this data is stored in the Anis database.
